Investigating social change during cultural contact period using geometric morphometry of pottery shapes from Iron Age northeastern Taiwan

### 2019-12-24 analysis/paper/paper.Rmd
updates for CV test

- CV is hard to test, we apply some statistics testing for it

# Materials and methods

@@

Pottery analysed in this paper comes from of Kiwulan, northeastern Taiwan. Kiwulan is situated on a hill near a riverside at the northern margin of Yilan, which is characterized by a triangular alluvial plain facing eastwards the Pacific and mountains on three other sides. Kiwulan can be divided into two cultural components, the upper and the lower component, with a sterile layer in between. This paper focus on the upper component dated to AD 1350 to AD 1950, which covers around 600 years from the late Iron Age to the historical period started with the European presence in Taiwan in the early 17th century. The Dutch first came and occupied southern Taiwan in 1624 and then the Spanish occupied northern Taiwan in 1626. In 1642, the Spanish was expelled by the Dutch, who took over their forts in northern Taiwan. Since then, western Taiwan was mostly under the colonial rule of the Dutch until 1662 when the Kingdom of Tungning was founded by Koxinga, a loyalist of Ming dynasty of China.

@@

## Outline analysis

@@

# statistical test

pc_plot_tests <- 
  pc_plot %>% 
  group_by_at(vars(-pc)) %>% 
  mutate(i = row_number()) %>% 
  ungroup() %>% 
  mutate(phase = as.character(phase),
         pcn = as.character(pcn)) %>% 
  spread(phase, pc) %>% 
  select(-i) %>% 
  nest_legacy(-pcn) %>% 
  mutate(var_test = purrr::map(data, ~broom::tidy(var.test(.x$`Post-European` ,
                                            .x$`Pre-European`)))) %>% 
  unnest(var_test)

@@

The distributions of the first three PCs for each vessel suggest variations in shape standardization by period. The first PC, the hight of the vessels, shows more variations in the pre-Euro period compared to the Chinese period. Standardization appears strong in the pre-Euro period but weak in the Chinese period. P-values for a modified signed-likelihood ratio test of equality of CVs of PC1 & PC2 show significant differences in standardization of vessel shape across some periods, especially between Chinese contact with either pre-European or post-European.

>>>>

- craft or pottery specialization
- CV is hard to test, we apply some statistics testing for it

# Materials and historical background

@@

Pottery analysed in this paper comes from Kiwulan, northeastern Taiwan. Kiwulan is situated on a hill near a riverside at the northern margin of Yilan, which is characterized by a triangular alluvial plain facing eastwards the Pacific and mountains on three other sides. Kiwulan can be divided into two cultural components, the upper and the lower component, with a sterile layer in between. This paper focus on the upper component dated to AD 1350 to AD 1950, which covers around 600 years from the late Iron Age to the historical period started with the European presence in Taiwan in the early 17th century. The Dutch first came and occupied southern Taiwan in 1624 and then the Spanish occupied northern Taiwan in 1626. In 1642, the Spanish was expelled by the Dutch, who took over their forts in northern Taiwan. Since then, western Taiwan was mostly under the colonial rule of the Dutch until 1662 when the Kingdom of Tungning was founded by Koxinga, a loyalist of Ming dynasty of China.

@@

# Methods

@@

# statistical test
pc_plot_tests <- 
  pc_plot %>% 
  group_by_at(vars(-pc)) %>% 
  mutate(i = row_number()) %>% 
  ungroup() %>% 
  mutate(phase = as.character(phase),
         pcn = as.character(pcn)) %>% 
  spread(phase, pc) %>% 
  select(-i) %>% 
  nest_legacy(-pcn) %>% 
  mutate(var_test = purrr::map(data, ~broom::tidy(var.test(.x$`Post-European` ,
                                            .x$`Pre-European`)))) %>% 
  unnest(var_test)

@@

The standardization of pottery shape for each phase is explored by the distributions of the first three PC scores for each vessel that suggests variations in shape standardization by phase. The first PC, the hight of the vessels, shows more variations in the pre-Euro period compared to the Chinese period. That shows standardization appears strong in the Chinese period but weak in the pre-European period. To see whether the differences of the distribution of PCs between each phase is statistically significant, coefficients of variation statistic (CVs) among phase groups of PCs is examined with a modified signed-likelihood ratio test for the equality of CVs. P-values for the significant test of equality of CVs of PC1 and PC2 show significant differences in standardization of vessel shape across some periods, especially between Chinese contact with either pre-European or post-European.
